\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}
I want to express my gratitude to Prof. Dr. Aggelos Kiayias for his supervision. I want to thank explicitly Dionysis Zindros, for his crucial assistance in the progress of this thesis, his support and his patience. I highlight the help I got from the thesis of Kostis Karantias and his work. Christos Nasikas, colleague and friend, has offered academic and personal guidance every time I needed it.

Along with Prof. Dr. Aggelos Kiayias, I am also thankful to the other two members of my three-member committee, Prof. Dr. Aris Pagourtzis and Prof. Dr. Nikolaos Papaspyrou, for the evaluation of my thesis.

During my master curriculum I had a huge offer of valuable courses and inspiring professors. This program has been an important factor of my academic and personal development.

My research was accompanied by a lot of interesting discussions and help from friends that took the time to listen about my thoughts and interests. Among them I should not neglect to name Yanna Papadodimitraki for her valuable contribution and Vassilis Agiotis for the illustration of my thesis. I thank for the support and patience my friends G.K., M.G., K.A., M.P. and C.V..

I also thank the Monero project and all of the people who decide to work in projects that have significant political and social impact. They have been an example for me.

For his invalueable contribution to my personality and culture development as an active member of society I have to thank my friend and my role model C.C.

Last, but most significantly, I want to salute all these people who do (or did) not have the chance to study and those who chose not to, in order to fight for freedom.
%
